\textbf{Response Letter}

### 2018-12-26 doc/aij18/major-revision/letter.tex
AIJ final ahora si que si

\textcolor{blue}{In the following, our responses to the reviewers are shown up in blue color.}

\vspace{0.3cm}

\textcolor{red}{Additionally, all the changes are highlighted in red in the document file.}

@@

\textcolor{blue}{We agree with the reviewer that the paper lacks a description of the overall approach. We refer the reviewer to the main modification \#1 exposed above. Following the notes of the reviewer's summary, we have also included a brief summary at the end of the section Introduction that serves to wrap up the main features of \FAMA. Additionally, and as requested by the other reviewers too, we introduced the main modification \#2. The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace in section 3.1. We believe that this explanation also helps to grasp the overall picture of \FAMA.}

@@

\textcolor{blue}{As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop. The author of LOUGA promptly provided us with the code, and so we thank him for his gentle response. The issue we faced though with LOUGA is that this system is built as a graphical interface application, which makes it impossible to automate the model learning process and evaluation process via a script for the 154 tested cases. Nevertheless, we did manually test LOUGA on a few domains, namely the \emph{blocksworld}, \emph{floor-tile}, \emph{n-puzzle} and \emph{transport} domains. We evaluated LOUGA and \FAMA in the experiment with FO action sequences and PO state trajectories with 10\% of observability using 5 traces. The global precision and recall values for each domain are shown below as LOUGA/\FAMA: }

\begin{table}[h!]
\begin{center}
\begin{tabular}{lll}
        & Precision  &  Recall    \\
 blocks &  0.78/1.0 &  0.78/0.93    \\
 floor-tile & 0.83/0.75  &  1.0/1.0 \\
 n-puzzle & 0.88/1.0 & 1.0/1.0  \\
 transport &  0.70/0.79 &  0.8/0.95  \\
\textcolor[rgb]{0.00,0.00,1.00}{AVERAGE}  & \textcolor{blue}{0.82/0.89} & \textcolor{blue}{0.90/0.97}
\end{tabular}
\end{center}
\end{table}

@@

\textcolor{blue}{The reviewer is right in that the definition of effects was confusing. It all stems from the adoption of the open world assumption, which leads to states that contain positive and negative literals, but we did not adapt the definition of effects appropriately. In section 2.1, we have now included a brief explanation of the open world assumption and the representation of states, and we have adapted the definition of effects accordingly.}

@@

\textcolor{blue}{The additional assumptions are the syntactic constrains imposed by \strips\ planning models. We agree with the reviewer that other domain models do not satisfy these assumptions. We intend to extend the compilation scheme to planning models other than \strips\ as future work, but the focus of this paper is exclusively on \strips\ models. We have included a sentence before the description of the assumptions to clearly state that they derive from the syntactic constraints of \strips.}

@@

\textcolor{blue}{We have reformulated the explanation about the observability in plan traces of section 2.1 to make it more comprehensible and added a few more details that help distinguish the different cases of observability. We have also highlighted the two special cases, PO* and NO, that derive from PO. With respect to the relation between actions and state partial observability, we refer the reviewer to the main modification \#2. The new section 3.1 presents the combinations of observed state trajectories and
observed action sequences and the impact in the learning task solving. Particularly, as it is written in the text at the end of page 10, the case FO actions with FO states is straightforward because in this case every state is available and so preconditions and effects are derivable by lifting the corresponding literals. The case FO actions with PO* states is a common case handled by some learning approaches as shown in Table 2 of section 2.2. In general, all learning approaches but \FAMA assume FO action sequences, which may come along with PO*, PO or NO state trajectories.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{Done. LOUGA is included in Table 2 and Table 3 and an explanatory paragraph about the main characteristics of LOUGA has been incorporated in section 2.2.}

@@

\textcolor{blue}{It depends on the domain we are trying to learn. Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.}

@@

\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects. This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn. Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}

\textcolor{blue}{On the other hand, the assumptions regarding the syntactic \strips\ constraints still apply when using conditional effects}

@@

\textcolor{blue}{We refer the reviewer to the main modification \#3. Section 4 has been polished significantly. An introductory paragraph provides a high level description of the approach and examples have been incorporated throughout the text.}

@@

\textcolor{blue}{Removing symmetries would admittedly improve performance for heuristic planners. This removal can be achieved by ensuring the actions are programmed following some order, which can be done via some additional fluents. We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.}

\textcolor{blue}{Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.}

@@

\textcolor{blue}{The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4. The caption of the figure indicates that "implications are coded as disjunctions". The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$.}

@@

\textcolor{blue}{The positions of the validate actions among the apply actions is determined by the planner, which will verify that the state that results after the execution of an apply action comprises the observed state of the plan trace. We have added the following sentence in the document to clarify this point: "There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$. The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."}

@@

\textcolor{blue}{All plotted values are averages over the 14 (now 15) tested domains. The reviewer is right, however, to point out that this is not said in the experiment of section 6.2, so we have added some lines making it explicit.}

@@

\textcolor{blue}{Following the reviewer's advice, we have completely remodeled section 3 and we have introduced major improvements in section 4. We refer the reviewer to our main modifications \#2 and \#3. We removed the complexity table of page 8, focusing now the discussion around the two main cases that determine the difficulty of the learning task and giving an indication of the task complexity according to the characteristics that define these two cases. We also changed the reference and added instead the original work that proves that boolean SAT problems are NP-complete.}

\textcolor{blue}{The take-home lesson of section 3 is that when the observed plan trace does not provide information of the solution plan length (case 2), planning is needed to find the solution and so the learning task is PSPACE-complete.}

@@

\textcolor{blue}{We have removed repetitions as much as possible. In the first example pointed by the reviewer, the explanation about partial observability of \FAMA and the other learning systems is concentrated in a single paragraph and the repetition is eliminated. Nevertheless, we think that at some points it is necessary to recall a particular concept in order to enhance the exposition of a further concept.}

\textcolor{blue}{With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$. We have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}

\textcolor{blue}{With respect to the third observation, we have introduced several changes in section 5. First, we provide the standard definitions of precision and recall at the beginning of the section. Then, we adapt these two concepts to planning problems in Def. 4 and 5, respectively. Subsequently, in section 5.2, we present our novel semantic-based metrics that build upon definitions 4 and 5.}

@@

\textcolor{blue}{Timeouts are only produced in the setting with NO action sequences and NO state trajectories. We have added some details on the domains where timeouts are found: in the grid domain when $|\mathcal{T}| = 3$, floortile when $|\mathcal{T}| = 4$ and four other domains. We refer the reviewer to the comment added in page 23 of section 6.2.}

@@

\textcolor{blue}{The reviewer is totally right in that planning is needed when the length of the solution plan is unknown. Precisely, this is the point that motivates our approach but that we did not successfully convey in the previous version of the paper. When the observed plan trace comprises a partially observed action sequence and a partially observed state trajectory, we ignore the length of the solution plan and so we need to resort to planning. The new section 3 explains this in detail (main modification \#2). When the length of the plan is known, SAT-solvers can be used; when it is not known, planning is required. In our case, we use a planner to solve all learning tasks independently of their nature.}

\textcolor{blue}{On the other hand, we happen to use a SAT-based planner (Madagascar) for solving the learning tasks, which may lead to wrongly interpret that the length of the solution plan is always known. The reasons for using Madagascar have been highlighted in red at the beginning of section 6. In a nutshell, we use Madagascar because it returned the best results compared to other planners and because of its ability to handle problems with dead-ends. Additionally, Madagascar is obviously very suitable to solve learning tasks when the length of the plan is known.}

@@

\textcolor{blue}{We acknowledge the reviewer's concern and refer the reviewer to the main modification \#1 exposed at the beginning of this response letter.}
\vspace{0.1cm}

@@

\textcolor{blue}{We have replaced the word 'abstract' by 'approximation' and rephrased the sentence a little bit.}
\vspace{0.1cm}

@@

\textcolor{blue}{We think this paragraph was rather confusing and that we did not clearly come through our intention to set an analogy between the data sources used in explainable AI and the data required by the learning approaches. For this reason, we have removed the paragraph and included instead a list of planning-related tasks that require a planning model for being solved. We think this rather complete but non-exhaustive list is a pretty good indication of the need for learning planning models.}

@@

\textcolor{blue}{We have rephrased all the paragraphs and sentences pointed by the reviewer. With respect to the first comment, we attempt to make it clear that classical planning techniques can be used to generate a large variety of planning models, not only \strips\ models, and that these techniques underpin some relevant concepts that are also reusable in learning models when the behaviour of the agent is not available.}

\textcolor{blue}{With respect to the other two comments, besides using different expressions to refer to the two concepts, we have included some additional side explanations that we hope help understand the meaning of the sentences.}
\vspace{0.1cm}

@@

\textcolor{blue}{We have included an explanation of the open world assumption and how this affects the representation of states in Section 2.1.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{Fixed.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{The sentence was, indeed, unclear. Anyway, we removed it in the revised version of the paper since it was an expendable phrase that did not fit very well in this section either.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{We have moved the 4th paragraph of the Related Work, where we commented on the predicates and action headers of the approaches, to the end of the section since it makes more sense to comment on this issue when comparing with LOCM. Also, we clarified the meaning of action header (action name and parameters) because the formal definition of an action header is actually introduced in section 4.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{The labels FO, PO*, PO and NO concern exclusively the intermediate states of the state trajectories as indicated in Table 1. We have improved the explanation of the four types of state trajectories in section 2.1. As for LOCM, this approach does not require any state, not even the initial or final state. We changed the label of LOCM in the first column of Table 2 to "---" to indicate that no state information at all is needed. We have also rephrased the sentence about LOCM.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{The reviewer is right in that \FAMA also assumes that predicates and action headers are provided or otherwise are extractable from the plan traces. We believe it is more appropriate to move this comment after LOCM is introduced since LOCM is an exception that learns the predicates using Finite-State Machines (FSM) because no observed states are given in the plan trace.}

@@

\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}

@@

\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}

@@

\textcolor[rgb]{0.00,0.07,1.00}{We refer the reviewer to the main modification \#2. Section 3 has substantially changed and this sentence does no longer appear in the text.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{It is true that a machinery for capturing and labeling the executed actions can be deployed. In either case, whether a human or an automated mechanism, the need of a fully observable behaviour of the agent imposes a limitation in learning systems. We have soften the sentence including the word "\emph{commonly}" in our exposition.}

@@

\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}

@@

\textcolor[rgb]{0.00,0.07,1.00}{We have removed the paragraph regarding the evaluation of the learned models from section 3 and moved to section 5. We also clarify the use of the word "semantic" as requested by another reviewer as well.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{In the remodeling of section 3, the paragraph about the size of the training dataset has been removed as well. The ideas presented here have been moved to page 23, at the end of section 6.2.}

@@

\textcolor{blue}{The function $ar(\cdot)$ does not return a set, it returns the arity of an action or predicate; i.e., the length of its argument list. This is introduced in the Background section for the first time.}

@@

\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}

@@

\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects. This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn. Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}

@@

\textcolor[rgb]{0.00,0.07,1.00}{We have introduced a new paragraph at the beginning of section 4 that gives an overview of the compilation approach and explains the two blocks of actions of the solution plan to the learning task. We have reused the concept of "block of actions" to refer to the prefix and postfix of $\pi_{\Lambda}$ in section 4.2 "Compilation".}

@@

\textcolor[rgb]{0.00,0.07,1.00}{$\Rightarrow$ is the general logic symbol read as implication. The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4. The caption of the figure indicates that "implications are coded as disjunctions". The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$ }

@@

\textcolor{blue}{We have added a new sentence that overviews the post-process. We have also rephrases some part of the text to make it more comprehensible. We believe the text provides now a clear idea of the post-process and that the pseudo-code is not necessary. Nevertheless, we refer the reviewer to the pseudo-code presented in [40] (algorithm 2) for a more detailed explanation.}

@@

\textcolor{blue}{We have provided better justification for the use of semantic evaluation metrics in section 5.}

@@

\textcolor{blue}{The precision and recall metrics give a notion of the soundness and completeness of a model. By a sound model we understand one where all preconditions and effects are correct, either wrt to the GTM or a test set, so there is no need to remove anything. A complete model is one where no precondition or effect is missing. We have added a sentence to clarify this.}

@@

\textcolor{blue}{Fixed.}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
Pages 15-19.
-  "... the horizon of the solution plan is also known. " - how do you know the horizon? You know the length of the given traces, but how do you know how many actions are needed to build the action model?
\end{mdframed}

\textcolor{blue}{Hopefully the new section 3 (see main modification \#2) helps clarify this question. In a nutshell, sometimes a plan trace indicates the length of the plan (this is the case when SAT solvers can be used) and others it does not (this is the case when planning is required).}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  "... a solution plan is solvable in two steps" - which two steps? I thought you put it all in the planner and get both action model and validation in the same step.
\end{mdframed}

\textcolor{blue}{Yes, the reviewer is right in that we put it all in the planner and get the overall plan, both building actions and validating actions. What we mean by "solvable in two steps" is that a SAT-based planner internally solves the construction of the model in two \emph{time steps}: in one time step the truth value of all the model fluents that represent the preconditions is set, and in a second time step the value of all the model fluents that represent the effects is set. We have slightly modified the sentence to make it clear we refer to the SAT solving process of Madagascar.}

@@

\textcolor{blue}{Fixed.}

@@

\textcolor{blue}{Yes, precision gives a notion of soundness since it measures how many of the learned preconditions and effects are correct. We have improved the explanation of the metrics in section 5 to make this more clear.}

@@

\textcolor{blue}{We have replaced the word "complete" by "adequate" as suggested by the reviewer.}

@@

\textcolor{blue}{We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}. In our claim of \FAMA using the minimal amount of information, we should actually say that \FAMA uses the \emph{minimal observability} possible, being the extreme case to have no actions at all and only the initial state and final state as observations. In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed.}

\textcolor{blue}{We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table 2 and, in this regard, we have highlighted this distinction in the explanation of section 2.2. Besides adding a sentence that stresses the ability of LOCM to learning with minimal input knowledge, we devote the last paragraph of section 2.2. to explain the distinctive features of the LOCM family.}

@@

\textcolor{blue}{The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects. This gives rise to an optimistic measure of the error since the set of possible preconditions and effects is usually much larger than the size of the actual GTM, resulting in low error rates even when the model is largely flawed. In contrast, precision and recall normalize by the size of the GTM, thus removing this problem. Furthermore, precision and recall provide an intuitive notion of soundness and completeness, respectively; whereas a sound model is interpreted as a model in which all preconditions and effects are correct wrt the GTM, a complete model is one in which no precondition or effect is missing. We have added some paragraphs to section 5 to better explain these concepts.}

>>>>

\textcolor{red}{All the changes are highlighted in red in the document file.}

@@

We agree with the reviewer that the paper lacks a description of the overall approach. We refer the reviewer to the main modification \#1 exposed above. Following the notes of the reviewer's summary, we have also included a brief summary at the end of the section Introduction that serves to wrap up the main features of \FAMA. Additionally, and as requested by the other reviewers too, we introduced the main modification \#2. The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace in section 3.1. We believe that this explanation also helps to grasp the overall picture of \FAMA.

@@

As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop. The author of LOUGA promptly provided us with the code, and so we thank him for his gentle response. The issue we faced though with LOUGA is that this system is built as a graphical interface application, which makes it impossible to automate the model learning process and evaluation process via a script for the 154 tested cases. Nevertheless, we did manually test LOUGA on a few domains, namely the \emph{blocksworld}, \emph{floor-tile}, \emph{n-puzzle} and \emph{transport} domains. We evaluated LOUGA and \FAMA in the experiment with FO action sequences and PO state trajectories with 10\% of observability using 5 traces. The global precision and recall values for each domain are shown below as LOUGA/\FAMA:

\begin{table}[h!]
\begin{center}
\begin{tabular}{lll}
        & Precision  &  Recall    \\
 blocks &  0.78/1.0 &  0.78/0.93    \\
 floor-tile & 0.83/0.75  &  1.0/1.0 \\
 n-puzzle & 0.88/1.0 & 1.0/1.0  \\
 transport &  0.70/0.79 &  0.8/0.95  \\
AVERAGE  & 0.82/0.89 & 0.90/0.97
\end{tabular}
\end{center}
\end{table}

@@

The reviewer is right in that the definition of effects was confusing. It all stems from the adoption of the open world assumption, which leads to states that contain positive and negative literals, but we did not adapt the definition of effects appropriately. In section 2.1, we have now included a brief explanation of the open world assumption and the representation of states, and we have adapted the definition of effects accordingly.

@@

The additional assumptions are the syntactic constrains imposed by \strips\ planning models. We agree with the reviewer that other domain models do not satisfy these assumptions. We intend to extend the compilation scheme to planning models other than \strips\ as future work, but the focus of this paper is exclusively on \strips\ models. We have included a sentence before the description of the assumptions to clearly state that they derive from the syntactic constraints of \strips.

@@

We have reformulated the explanation about the observability in plan traces of section 2.1 to make it more comprehensible and added a few more details that help distinguish the different cases of observability. We have also highlighted the two special cases, PO* and NO, that derive from PO. With respect to the relation between actions and state partial observability, we refer the reviewer to the main modification \#2. The new section 3.1 presents the combinations of observed state trajectories and
observed action sequences and the impact in the learning task solving. Particularly, as it is written in the text at the end of page 10, the case FO actions with FO states is straightforward because in this case every state is available and so preconditions and effects are derivable by lifting the corresponding literals. The case FO actions with PO* states is a common case handled by some learning approaches as shown in Table 2 of section 2.2. In general, all learning approaches but \FAMA assume FO action sequences, which may come along with PO*, PO or NO state trajectories.

@@

Done. LOUGA is included in Table 2 and Table 3 and an explanatory paragraph about the main characteristics of LOUGA has been incorporated in section 2.2.

@@

It depends on the domain we are trying to learn. Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.

@@

Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects. This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn. Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.

On the other hand, the assumptions regarding the syntactic \strips\ constraints still apply when using conditional effects

@@

We refer the reviewer to the main modification \#3. Section 4 has been polished significantly. An introductory paragraph provides a high level description of the approach and examples have been incorporated throughout the text.

@@

Removing symmetries would admittedly improve performance for heuristic planners. This removal can be achieved by ensuring the actions are programmed following some order, which can be done via some additional fluents. We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.

Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.

@@

The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4. The caption of the figure indicates that "implications are coded as disjunctions". The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$.

@@

The positions of the validate actions among the apply actions is determined by the planner, which will verify that the state that results after the execution of an apply action comprises the observed state of the plan trace. We have added the following sentence in the document to clarify this point: "There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$. The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."

@@

All plotted values are averages over the 14 (now 15) tested domains. The reviewer is right, however, to point out that this is not said in the experiment of section 6.2, so we have added some lines making it explicit.

@@

Following the reviewer's advice, we have completely remodeled section 3 and we have introduced major improvements in section 4. We refer the reviewer to our main modifications \#2 and \#3. We removed the complexity table of page 8, focusing now the discussion around the two main cases that determine the difficulty of the learning task and giving an indication of the task complexity according to the characteristics that define these two cases. We also changed the reference and added instead the original work that proves that boolean SAT problems are NP-complete.

The take-home lesson of section 3 is that when the observed plan trace does not provide information of the solution plan length (case 2), planning is needed to find the solution and so the learning task is PSPACE-complete.

@@

We have removed repetitions as much as possible. In the first example pointed by the reviewer, the explanation about partial observability of \FAMA and the other learning systems is concentrated in a single paragraph and the repetition is eliminated. Nevertheless, we think that at some points it is necessary to recall a particular concept in order to enhance the exposition of a further concept.

With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$. We have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.

With respect to the third observation, we have introduced several changes in section 5. First, we provide the standard definitions of precision and recall at the beginning of the section. Then, we adapt these two concepts to planning problems in Def. 4 and 5, respectively. Subsequently, in section 5.2, we present our novel semantic-based metrics that build upon definitions 4 and 5.

@@

Timeouts are only produced in the setting with NO action sequences and NO state trajectories. We have added some details on the domains where timeouts are found: in the grid domain when $|\mathcal{T}| = 3$, floortile when $|\mathcal{T}| = 4$ and four other domains. We refer the reviewer to the comment added in page 23 of section 6.2.

@@

The reviewer is totally right in that planning is needed when the length of the solution plan is unknown. Precisely, this is the point that motivates our approach but that we did not successfully convey in the previous version of the paper. When the observed plan trace comprises a partially observed action sequence and a partially observed state trajectory, we ignore the length of the solution plan and so we need to resort to planning. The new section 3 explains this in detail (main modification \#2). When the length of the plan is known, SAT-solvers can be used; when it is not known, planning is required. In our case, we use a planner to solve all learning tasks independently of their nature.

On the other hand, we happen to use a SAT-based planner (Madagascar) for solving the learning tasks, which may lead to wrongly interpret that the length of the solution plan is always known. The reasons for using Madagascar have been highlighted in red at the beginning of section 6. In a nutshell, we use Madagascar because it returned the best results compared to other planners and because of its ability to handle problems with dead-ends. Additionally, Madagascar is obviously very suitable to solve learning tasks when the length of the plan is known.

@@

We acknowledge the reviewer's concern and refer the reviewer to the main modification \#1 exposed at the beginning of this response letter.
\vspace{0.1cm}

@@

We have replaced the word 'abstract' by 'approximation' and rephrased the sentence a little bit.

@@

We think this paragraph was rather confusing and that we did not clearly come through our intention to set an analogy between the data sources used in explainable AI and the data required by the learning approaches. For this reason, we have removed the paragraph and included instead a list of planning-related tasks that require a planning model for being solved. We think this rather complete but non-exhaustive list is a pretty good indication of the need for learning planning models.

@@

We have rephrased all the paragraphs and sentences pointed by the reviewer. With respect to the first comment, we attempt to make it clear that classical planning techniques can be used to generate a large variety of planning models, not only \strips\ models, and that these techniques underpin some relevant concepts that are also reusable in learning models when the behaviour of the agent is not available.

With respect to the other two comments, besides using different expressions to refer to the two concepts, we have included some additional side explanations that we hope help understand the meaning of the sentences.

@@

We have included an explanation of the open world assumption and how this affects the representation of states in Section 2.1.

@@

Fixed.

@@

The sentence was, indeed, unclear. Anyway, we removed it in the revised version of the paper since it was an expendable phrase that did not fit very well in this section either.

@@

We have moved the 4th paragraph of the Related Work, where we commented on the predicates and action headers of the approaches, to the end of the section since it makes more sense to comment on this issue when comparing with LOCM. Also, we clarified the meaning of action header (action name and parameters) because the formal definition of an action header is actually introduced in section 4.

@@

The labels FO, PO*, PO and NO concern exclusively the intermediate states of the state trajectories as indicated in Table 1. We have improved the explanation of the four types of state trajectories in section 2.1. As for LOCM, this approach does not require any state, not even the initial or final state. We changed the label of LOCM in the first column of Table 2 to "---" to indicate that no state information at all is needed. We have also rephrased the sentence about LOCM.

@@

The reviewer is right in that \FAMA also assumes that predicates and action headers are provided or otherwise are extractable from the plan traces. We believe it is more appropriate to move this comment after LOCM is introduced since LOCM is an exception that learns the predicates using Finite-State Machines (FSM) because no observed states are given in the plan trace.

@@

The figures that appear in the third column of Table 3 for each learning system were taken from the respective published papers. They all refer to the specific experimental setup used in the evaluation of the approaches. In general, since there is not a common benchmark for the evaluation of learning approaches, each system selects the configuration of training data that best suits the particularities of the approach. Hence, all the figures in the 3rd column of Table 3 refer to the specific implementation used in the experiments. What we tried to do in this section is to somehow compare the approaches by establishing some common elements of evaluation, namely number of tested domains, size of the training data in terms of action and/or traces, degree of observability, etc.

@@

A sort is a group of objects found in the same position in the header of an action. Although this was already defined in the text, we have changed the font of the word to italics and we explicitly say now this set is "named \emph{sort"}.

@@

We refer the reviewer to the main modification \#2. Section 3 has substantially changed and this sentence does no longer appear in the text.

@@

Corrected.

@@

It is true that a machinery for capturing and labeling the executed actions can be deployed. In either case, whether a human or an automated mechanism, the need of a fully observable behaviour of the agent imposes a limitation in learning systems. We have soften the sentence including the word "\emph{commonly}" in our exposition.

@@

We agree with the reviewer that relevant results have recently been achieved in the setting of full observability. Now we acknowledge, in Section 3.1. of our paper, the work by Stern and Juba that allows computing the least number of examples that are necessary to learn models within a given error rate.

@@

We have removed the paragraph regarding the evaluation of the learned models from section 3 and moved to section 5. We also clarify the use of the word "semantic" as requested by another reviewer as well.

@@

In the remodeling of section 3, the paragraph about the size of the training dataset has been removed as well. The ideas presented here have been moved to page 23, at the end of section 6.2.

@@

The function $ar(\cdot)$ does not return a set, it returns the arity of an action or predicate; i.e., the length of its argument list. This is introduced in the Background section for the first time.

@@

The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ because the value of two bits is enough to code how a given element from $\Psi_{\xi}$ appears in a given action model $\xi$: (0,0) indicates it does not appear in $\xi$, (1,0) indicates it appears as precondition, (1,1) indicates it appears as precondition and a negative effect and (0,1) indicates it appears as a positive effect.

@@

Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects. This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn. Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.

@@

We have introduced a new paragraph at the beginning of section 4 that gives an overview of the compilation approach and explains the two blocks of actions of the solution plan to the learning task. We have reused the concept of "block of actions" to refer to the prefix and postfix of $\pi_{\Lambda}$ in section 4.2 "Compilation".

@@

$\Rightarrow$ is the general logic symbol read as implication. The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4. The caption of the figure indicates that "implications are coded as disjunctions". The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$

@@

We have added a new sentence that overviews the post-process. We have also rephrased some part of the text to make it more comprehensible. We believe the text provides now a clear idea of the post-process. Nevertheless, we refer the reviewer to the pseudo-code presented in [40] (algorithm 2) for a more detailed explanation.

@@

We have provided better justification for the use of semantic evaluation metrics in section 5.

@@

The precision and recall metrics give a notion of the soundness and completeness of a model. By a sound model we understand one where all preconditions and effects are correct, either wrt to the GTM or a test set, so there is no need to remove anything. A complete model is one where no precondition or effect is missing. We have added a sentence to clarify this.

@@

Fixed.

-  "... the horizon of the solution plan is also known. " - how do you know the horizon? You know the length of the given traces, but how do you know how many actions are needed to build the action model?

-  "... a solution plan is solvable in two steps" - which two steps? I thought you put it all in the planner and get both action model and validation in the same step.
\end{mdframed}

Yes, the reviewer is right in that we put it all in the planner and get the overall plan, both building actions and validating actions. What we mean by "solvable in two steps" is that a SAT-based planner internally solves the construction of the model in two \emph{time steps}: in one time step the truth value of all the model fluents that represent the preconditions is set, and in a second time step the value of all the model fluents that represent the effects is set.

%We have slightly modified the sentence to make it clear we refer to the SAT solving process of Madagascar.

@@

Fixed.

@@

Yes, precision gives a notion of soundness since it measures how many of the learned preconditions and effects are correct. We have improved the explanation of the metrics in section 5 to make this more clear.

@@

We have replaced the word "complete" by "adequate" as suggested by the reviewer.

@@

We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}. In our claim of \FAMA using the minimal amount of information, we should actually say that \FAMA uses the \emph{minimal observability} possible, being the extreme case to have no actions at all and only the initial state and final state as observations. In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed. We want to point out, however, that \FAMA does not require the number of steps that are in the training plan, and we hope the new section 3 (see main modification \#2) does a better job at conveying this point.

We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table 2 and, in this regard, we have highlighted this distinction in the explanation of section 2.2. Besides adding a sentence that stresses the ability of LOCM to learning with minimal input knowledge, we devote the last paragraph of section 2.2. to explain the distinctive features of the LOCM family.

@@

The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects. This gives rise to an optimistic measure of the error since the set of possible preconditions and effects is usually much larger than the size of the actual GTM, resulting in low error rates even when the model is largely flawed. In contrast, precision and recall normalize by the size of the GTM, thus removing this problem. Furthermore, precision and recall provide an intuitive notion of soundness and completeness, respectively; whereas a sound model is interpreted as a model in which all preconditions and effects are correct wrt the GTM, a complete model is one in which no precondition or effect is missing. We have added some paragraphs to section 5 to better explain these concepts.

As the reviewer points out, \FAMA will create a model that explains all the training traces because it is a satisfying approach (that satisfies all the training samples) and so the training error is 0. There is, however, no guarantee that the training traces are able to capture the whole dynamics of the domain. This is specially true for scenarios with low observability like the ones explored in this paper. Hence, it is always interesting to measure how well the learned model is able to explain a new set of plan traces.
